---
title: "Class Activation Map"
output:
  html_document:
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float: false
    code_folding: show
editor_options: 
  chunk_output_type: console
---


> This demo shows the method proposed in [Learning Deep Features for Discriminative Localization](https://arxiv.org/abs/1512.04150)

This is an R port of the Python demo found [here](https://github.com/dmlc/mxnet-notebooks/blob/master/python/moved-from-mxnet/class_active_maps.ipynb).

Idea is to apply the last fully connected layer to each of the pixel following the last convolution layer. In Resnet, this applies to the 7x7 final features, just before the global average pooling. 

Since the final FC is meant to create a score for each of the 1000 classes, applying those FC weights to the 7x7features prior their pooling result in a class score for each of those pixel. 

```{r, echo=TRUE, message=FALSE}
require("grid")
require("readr")
require("dplyr")
require("plotly")
require("imager")
require("mxnet")
```


## Load model

A ResNet model is loaded and its last convolution output is grouped together the softmax output. Also, the weights of the last fully connected output is extracted. 

These are the key ingredients of the class activation map. 

```{r, echo=TRUE}
resnet <- mx.model.load("../models/resnet-50", iteration=0)

labels_resnet <- readLines("../models/synset.txt")
labels_resnet <- gsub(pattern = "(n\\d+\\s)(.+)", replacement="\\2", x=labels_resnet)

symbol <- resnet$symbol
internals <- symbol$get.internals()
outputs <- internals$outputs

# last layer before global pooling
conv <- internals$get.output(which(outputs=="relu1_output"))
# flatten layer
softmax <- internals$get.output(which(outputs=="softmax_output"))
symbol_group <- mx.symbol.Group(c(conv, softmax))
# last fully connected weights
weight_fc <- resnet$arg.params$fc1_weight
```


## Image treatment

Add padding to image and reshape into a 224 X 224 as required by ResNet model. 

```{r}
preproc_resnet_pad <- function(im, resize=224) {
  shape <- dim(im)[1:2]
  axis = c("x", "y")[which.min(shape)]
  pad = abs(diff(shape))
  im <- pad(im, nPix=pad, axes = axis)
  resized <- resize(im, resize, resize)
  array <- as.array(resized) * 255
  dim(array) <- c(resize, resize, 3, 1)
  return(array)
}

im <- imager::load.image("../models/cat-dog-2.jpg")

data_array <- preproc_resnet_pad(im)
data = mx.nd.array(data_array)
label = mx.nd.array(1)

im_array <- data_array
dim(im_array) <- dim(im_array)[1:3]
plot(im_array %>% as.cimg())
```


## Generate activation map

```{r}
arg_names <- symbol_group$arguments
arg.arrays <- c(list(data=data, softmax_label = label), resnet$arg.params)[arg_names]
aux.arrays <- resnet$aux.params[symbol_group$auxiliary.states]

# outputs = mod.predict(blob)
exec <- mxnet:::mx.symbol.bind(symbol = symbol_group, ctx = mx.cpu(), 
                                 arg.arrays = arg.arrays, 
                                 aux.arrays = aux.arrays,
                                 grad.reqs = rep("null", length(arg.arrays)))

mx.exec.forward(exec, is.train = F)
outputs <- exec$outputs
score = outputs[[2]]
conv_fm = outputs[[1]]
dim(conv_fm)

# get the indices of the topk predictions
top_k = 8
inds_topk <- mx.nd.topk(data=score, axis=1, k=top_k) %>% mx.nd.reshape(-1)
labels_resnet[as.array(inds_topk)+1]
```

## CAM function

Apply the last fully connected operation to each of the final 7x7 features. 
Only the weights relevant to the top k classes are kept from the FC weights. 

```{r}
get_cam <- function(conv_fm, weight_fc) {
  
  conv_fm = mx.nd.reshape(data = conv_fm, shape = c(0,0,0), reverse = T)
  conv_fm_flatten = mx.nd.reshape(data = conv_fm, shape = c(-1,0))
  dim(conv_fm)
  dim(conv_fm_flatten)
  dim(weight_fc)

  # results in shape (height X width) X topk
  detection_map = mx.nd.dot(lhs = weight_fc, rhs = conv_fm_flatten)
  map_shapes <- dim(detection_map)
  detection_map = mx.nd.reshape(detection_map, shape=c(sqrt(map_shapes[1]), sqrt(map_shapes[1]), map_shapes[2]))
  return(detection_map)
}
```


## Plot activation map

Plot image with activation mask for the top 4 labels. 

```{r, echo = FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r, fig.height=12, fig.width=8}
weight_fc_topk = mx.nd.take(weight_fc, indices = inds_topk, axis = 0)
cam = get_cam(conv_fm = conv_fm, weight_fc = weight_fc_topk)
cam_array = cam %>% as.array()
ori_image = im_array %>% as.cimg()
ori_image = ori_image  / max(ori_image)

plots <- list()

for (k in 1:top_k) {
  cam_img = cam_array[, , k] %>% as.cimg()
  heat_map = resize(cam_img, size_x = dim(im_array)[1], size_y = dim(im_array)[1], interpolation_type = 3)
  max_response = mean(cam_img)
  heat_map_color = hsv(h = (heat_map - min(heat_map))/(max(heat_map) - min(heat_map))) %>% 
    col2rgb %>% t %>% as.vector %>% as.cimg(dim=c(dim(heat_map)[1:3],3))
  
  heat_map_color <- heat_map_color / 255
  im_comb <- imdraw(ori_image, heat_map_color, opacity = 0.5)
  
  df <- as.data.frame(im_comb,wide="c") %>% mutate(rgb.val=rgb(c.1,c.2,c.3))
  p <- ggplot(df,aes(x,y))+geom_raster(aes(fill=rgb.val))+scale_fill_identity()
  p <- p+scale_x_continuous(expand=c(0,0))+scale_y_continuous(expand=c(0,0),trans=scales::reverse_trans()) + ggtitle(labels_resnet[as.array(inds_topk)+1][k])+coord_fixed(ratio = 1) +
    theme(axis.title=element_blank())
  plots[[k]] <- p
}

multiplot(plotlist = plots, cols = 2)
```


