---
title: "CNN - NLP Classification"
output: github_document
editor_options: 
  chunk_output_type: console
---

> Demo the application of CNN for NLP task with varying length sequences using both padding and bucketing through custom iterator with MXNet R package

Example based on sentiment analysis on the [IMDB data](http://ai.stanford.edu/~amaas/data/sentiment/).

Load some packages

```{r, echo=FALSE, message=FALSE}
library("readr")
library("dplyr")
library("plotly")
library("stringr")
library("stringi")
library("AUC")
library("scales")
library("mxnet")
library("DiagrammeR")

source("../rnn.infer.R")
```

## Data preparation

For this demo, the data preparation is performed by the script `data_preprocessing_seq_to_one.R` which involves the following steps:

- Import IMDB movie reviews  
- Split each review into a word vector and apply some common cleansing (remove special characters, lower case, remove extra blank space...)  
- Convert words into integers and define a dictionary to map the resulting indices with former words  
- Aggregate the buckets of samples and labels into a list

- `corpus_single_train.rds`: no bucketing, all samples are padded/trimmed to 600 words.  

```{r, echo=FALSE, warning=FALSE, message=FALSE}
corpus_single_train <- readRDS(file = "../data/corpus_single_train.rds")
corpus_single_test <- readRDS(file = "../data/corpus_single_test.rds")

vocab <- length(corpus_single_train$dic)

### Create iterators
batch.size = 64

train.data.single <- mx.io.bucket.iter(buckets = corpus_single_train$buckets, batch.size = batch.size, 
                                       data.mask.element = 0, shuffle = TRUE)

eval.data.single <- mx.io.bucket.iter(buckets = corpus_single_test$buckets, batch.size = batch.size, 
                                      data.mask.element = 0, shuffle = FALSE)
```

## Model architecture

```{r, echo=TRUE, fig.height=2}
cnn.symbol.factory <- function(seq_len, 
                               input_size,
                               num_embed, 
                               num_filters,
                               num_decode, 
                               dropout,
                               ignore_label=-1) {
  
  conv_params <- list(embed_weight=mx.symbol.Variable("embed_weight"),
                      conv1.weight = mx.symbol.Variable("conv1_weight"),
                      conv1.bias = mx.symbol.Variable("conv1_bias"),
                      conv2.weight = mx.symbol.Variable("conv2_weight"),
                      conv2.bias = mx.symbol.Variable("conv2_bias"),
                      conv3.weight = mx.symbol.Variable("conv3_weight"),
                      conv3.bias = mx.symbol.Variable("conv3_bias"),
                      fc1.weight = mx.symbol.Variable("fc1_weight"),
                      fc1.bias = mx.symbol.Variable("fc1_bias"),
                      fc_final.weight = mx.symbol.Variable("fc_final.weight"),
                      fc_final.bias = mx.symbol.Variable("fc_final.bias"))
  
  # embeding layer
  label <- mx.symbol.Variable("label")
  data <- mx.symbol.Variable("data")
  data_mask <- mx.symbol.Variable("data_mask")
  data_mask_array <- mx.symbol.Variable("data_mask_array")
  data_mask_array<- mx.symbol.BlockGrad(data_mask_array)
  
  embed <- mx.symbol.Embedding(data=data, weight=conv_params$embed_weight, input_dim=input_size, output_dim=num_embed, name="embed")
  embed <- mx.symbol.swapaxes(data = embed, dim1 = 1, dim2 = 2)
  embed_expand <- mx.symbol.expand_dims(data=embed, axis=-1, name="embed_expand")
  
  conv1<- mx.symbol.Convolution(data=embed_expand, weight=conv_params$conv1.weight, bias=conv_params$conv1.bias, kernel=c(1, 3), stride=c(1,1), pad=c(0,1), num_filter=8)
  act1<- mx.symbol.Activation(data=conv1, act.type="relu", name="act1")
  pool1<- mx.symbol.Pooling(data=act1, global.pool=F, pool.type="avg" , kernel=c(1,2), stride=c(1,2), pad=c(0,0), name="pool1")
  
  conv2<- mx.symbol.Convolution(data=pool1, weight=conv_params$conv2.weight, bias=conv_params$conv2.bias, kernel=c(1,3), stride=c(1,1), pad=c(0,1), num_filter=16)
  act2<- mx.symbol.Activation(data=conv2, act.type="relu", name="act2")
  pool2<- mx.symbol.Pooling(data=act2, global.pool=F, pool.type="avg" , kernel=c(1,2), stride=c(1,2), pad=c(0,0), name="pool2")
  
  conv3 <- mx.symbol.Convolution(data=pool2, weight=conv_params$conv3.weight, 
                                 bias=conv_params$conv3.bias, kernel=c(1, 3), 
                                 stride=c(1,1), pad=c(0,1), num_filter=32)
  act3 <- mx.symbol.Activation(data=conv3, act.type="relu", name="act3")
  pool3 <- mx.symbol.Pooling(data=act3, global.pool=T, pool.type="max", kernel=c(1,1), name="pool3")
  
  flatten <- mx.symbol.Flatten(data=pool3, name="flatten")
  
  # fc1 <- mx.symbol.FullyConnected(data=flatten, weight = conv_params$fc1.weight, 
  #                                 bias=conv_params$fc1.bias, num.hidden=16, name="fc1")
  # act_fc <- mx.symbol.Activation(data=fc1, act.type="relu", name="act_fc")
  # dropout<- mx.symbol.Dropout(data=act_fc, p=dropout, name="drop")
  
  fc_final<- mx.symbol.FullyConnected(data=flatten, weight=conv_params$fc_final.weight, bias=conv_params$fc_final.bias, num.hidden=2, name="fc_final")
  
  # Removed the ignore label in softmax
  loss <- mx.symbol.SoftmaxOutput(data = fc_final, label = label, name="loss")
  return(loss)
  
}

num_decode=2
num_embed=2
num_filters=16
seq_len = 600

symbol_cnn <- cnn.symbol.factory(seq_len = seq_len, input_size = vocab, 
                                num_embed = num_embed, num_filters = num_filters, 
                                num_decode = num_decode, dropout = 0.5, ignore_label = -1)

graph.viz(symbol_cnn, shape=c(seq_len, batch.size), direction = "LR")
```


## Model training
```{r, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
devices <- mx.cpu()

initializer <- mx.init.Xavier(rnd_type = "gaussian", factor_type = "avg", magnitude = 2.5)

optimizer <- mx.opt.create("rmsprop", learning.rate = 1e-3, gamma1 = 0.95, gamma2 = 0.92, 
                           wd = 1e-4, clip_gradient = 5, rescale.grad=1/batch.size)

logger <- mx.metric.logger()
epoch.end.callback <- mx.callback.log.train.metric(period = 1, logger = logger)
batch.end.callback <- mx.callback.log.train.metric(period = 50)

system.time(
  model <- mx.model.buckets(symbol = symbol_cnn,
                            train.data = train.data.single, eval.data = eval.data.single,
                            num.round = 5, ctx = devices, verbose = FALSE,
                            metric = mx.metric.accuracy, optimizer = optimizer,  
                            initializer = initializer,
                            batch.end.callback = batch.end.callback, 
                            epoch.end.callback = epoch.end.callback)
)

mx.model.save(model, prefix = "../models/model_cnn_classif_single", iteration = 1)
```

```{r logger1, echo=FALSE, eval=TRUE, fig.height=4}
plot_ly(x = seq_len(length(logger$train)), y = logger$train, 
        type = "scatter", mode = "markers+lines", name = "train") %>% 
  add_trace(y = logger$eval, type = "scatter", mode = "markers+lines", name = "eval")
```


## Plot word embeddings

Word representation can be visualized by looking at the assigned weights in any of the embedding dimensions. Here, we look simultaneously at the two embeddings learnt in the LSTM model. 

```{r embed, echo=FALSE, eval = TRUE, fig.height=4}
model <- mx.model.load(prefix = "../models/model_cnn_classif_single", iteration = 1)

dic <- corpus_single_train$dic
rev_dic <- corpus_single_train$rev_dic

embeddings_weights <- t(as.array(model$arg.params$embed_weight))
embeddings <- data.frame(label=rev_dic, embeddings_weights)
plot_words <- c("terrible", "awesome", "great", "fantastic", "worst", "awful", "pain", "painful", "recommend", "not", "poor", "wonderful", "shame", "ok", "okay", "excellent", "worst", "adequate", "fair", "enjoy", "good", "like", "interesting", "beautiful", "wasn't", "wonderful", "lost", "laugh", "laughable", "unfortunately", "attempt")

embed_ids <- match(plot_words, embeddings$label)
embeddings_sample <- embeddings[embed_ids, ]

p <- ggplot(embeddings_sample, aes(x = X1, y = X2, label = label))
p + geom_text(check_overlap = T, color="navy") + theme_bw() +  theme(panel.grid=element_blank()) + 
  labs(x = "embed 1", y = "embed 2")
```

Since the model attempts to predict the sentiment, it's no surprise that the 2 dimensions into which each word is projected appear correlated with words' polarity. Positive words are associated with lower values along the first embedding (_great_, _excellent_), while the most negative words appear at the far right (_terrible_, _worst_). 

## Inference on test data

```{r, echo=TRUE}
ctx <- mx.cpu(0)
batch.size <- 64

corpus_single_test <- readRDS(file = "../data/corpus_single_test.rds")

test.data <- mx.io.bucket.iter(buckets = corpus_single_test$buckets, batch.size = batch.size, 
                               data.mask.element = 0, shuffle = FALSE)
```


```{r, echo=TRUE, eval=TRUE, warning=FALSE}

mx.symbol.bind = mxnet:::mx.symbol.bind
mx.nd.arraypacker = mxnet:::mx.nd.arraypacker
infer <- mx.infer.rnn(infer.data = test.data, model = model, ctx = ctx)

pred_raw <- t(as.array(infer))
pred <- max.col(pred_raw, tie = "first") - 1
label <- unlist(lapply(corpus_single_test$buckets, function(x) x$label))

acc <- sum(label == pred)/length(label)
roc <- roc(predictions = pred_raw[, 2], labels = factor(label))
auc <- auc(roc)
```

Accuracy: `r percent(acc)`  

AUC: `r signif(auc, 4)` 

