---
title: "Language Model on CPU"
output: github_document
editor_options: 
  chunk_output_type: inline
---

> This tutorial presents an example of application of one-to-one RNN applied to text generation using the reworked functionnalities in MXNet R package.  

Example based on [Obama's speech](http://data.mxnet.io/data/char_lstm.zip).

Load some packages

```{r, echo=T, message=F}
library("readr")
library("dplyr")
library("plotly")
library("stringr")
library("stringi")
library("scales")
library("mxnet")
source("../rnn.graph.R")
```

## Data preparation

Data preparation is performed by the script: `data_preprocessing_one_to_one.R`. 

The following steps are executed: 

- Import speach data as a single character string. 
- Remove non printable characters. 
- Split text into individual characters. 
- Group characters into sequences of a fixed length, each sequence being a sample to the model.

```{r, echo=TRUE}
corpus_bucketed_train <- readRDS(file = "../data/train_buckets_one_to_one.rds")
corpus_bucketed_test <- readRDS(file = "../data/eval_buckets_one_to_one.rds")

vocab <- length(corpus_bucketed_test$dic)

### Create iterators
batch.size = 32

train.data <- mx.io.bucket.iter(buckets = corpus_bucketed_train$buckets, 
                                batch.size = batch.size, 
                                data.mask.element = 0, shuffle = TRUE)

eval.data <- mx.io.bucket.iter(buckets = corpus_bucketed_test$buckets, 
                               batch.size = batch.size,
                               data.mask.element = 0, shuffle = FALSE)
```

## Model arhcitecture

A one-to-one model configuration is specified since for each character, we want to predict the next one. For a sequence of length 100, there are also 100 labels, corresponding the same sequence of characters but offset by a position of +1. 

```{r, echo=TRUE}
rnn_graph <- rnn.graph.unroll(seq_len = 2,
                              num_rnn_layer = 1, 
                              num_hidden = 96,
                              input_size = vocab,
                              num_embed = 64, 
                              num_decode = vocab,
                              masking = F, 
                              loss_output = "softmax",
                              dropout = 0.2, 
                              ignore_label = -1,
                              cell_type = "lstm",
                              output_last_state = F,
                              config = "one-to-one")
```

```{r, echo=FALSE, fig.height=12}
graph.viz(rnn_graph, type = "graph", direction = "TD", 
          shape=list(data = c(2, batch.size),
                     label = c(2, batch.size)))
```


## Fit a LSTM model

Unroll the RNN to the length of the input sequence.

```{r, echo=FALSE}
symbol <- rnn.graph.unroll(seq_len = as.integer(train.data$bucket.names),
                           num_rnn_layer = 2, 
                           num_hidden = 96,
                           input_size = vocab,
                           num_embed = 64, 
                           num_decode = vocab,
                           masking = F, 
                           loss_output = "softmax",
                           dropout = 0.2, 
                           ignore_label = -1,
                           cell_type = "lstm",
                           output_last_state = F,
                           config = "one-to-one")
```


```{r, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
ctx <- mx.cpu()

initializer <- mx.init.Xavier(rnd_type = "gaussian", factor_type = "avg", magnitude = 3)

optimizer <- mx.opt.create("adadelta", rho = 0.9, eps = 1e-5, wd = 1e-8,
                           clip_gradient = 5, rescale.grad = 1/batch.size)

logger <- mx.metric.logger()
epoch.end.callback <- mx.callback.log.train.metric(period = 1, logger = logger)
batch.end.callback <- mx.callback.log.train.metric(period = 50)

mx.metric.custom_nd <- function(name, feval) {
  init <- function() {
    c(0, 0)
  }
  update <- function(label, pred, state) {
    m <- feval(label, pred)
    state <- c(state[[1]] + 1, state[[2]] + m)
    return(state)
  }
  get <- function(state) {
    list(name=name, value=(state[[2]]/state[[1]]))
  }
  ret <- (list(init=init, update=update, get=get))
  class(ret) <- "mx.metric"
  return(ret)
}

mx.metric.Perplexity <- mx.metric.custom_nd("Perplexity", function(label, pred) {
  label <- mx.nd.reshape(label, shape = -1)
  label_probs <- as.array(mx.nd.choose.element.0index(pred, label))
  batch <- length(label_probs)
  NLL <- -sum(log(pmax(1e-15, as.array(label_probs)))) / batch
  Perplexity <- exp(NLL)
  return(Perplexity)
})

system.time(
  model <- mx.model.buckets(symbol = symbol,
                            train.data = train.data, eval.data = eval.data, 
                            num.round = 5, ctx = ctx, verbose = TRUE,
                            metric = mx.metric.Perplexity, 
                            initializer = initializer, optimizer = optimizer, 
                            batch.end.callback = NULL, 
                            epoch.end.callback = epoch.end.callback)
)

mx.model.save(model, prefix = "../models/model_one_to_one_lstm_cpu", iteration = 1)
```

```{r, echo=FALSE, fig.height=5}
plot_ly(x = seq_len(length(logger$train)), y = logger$train, type = "scatter", mode = "markers+lines", name = "train") %>% 
  add_trace(y = logger$eval, type = "scatter", mode = "markers+lines", name = "eval")
```


## Inference on test data

Setup inference data. Need to apply preprocessing to inference sequence and convert into a infer data iterator. 

The parameters `output_last_state` is set to `TRUE` in order to access the state of the RNN cells when performing inference. 

```{r, echo=TRUE}
ctx <- mx.cpu()
batch.size <- 1

corpus_bucketed_train <- readRDS(file = "../data/train_buckets_one_to_one.rds")
dic <- corpus_bucketed_train$dic
rev_dic <- corpus_bucketed_train$rev_dic

infer_raw <- c("The United States are")
infer_split <- dic[strsplit(infer_raw, '') %>% unlist]
infer_length <- length(infer_split)

symbol.infer.ini <- rnn.graph.unroll(seq_len = infer_length,
                                     num_rnn_layer = 2, 
                                     num_hidden = 96,
                                     input_size = vocab,
                                     num_embed = 64, 
                                     num_decode = vocab,
                                     masking = F, 
                                     loss_output = "softmax",
                                     dropout = 0.2, 
                                     ignore_label = -1,
                                     cell_type = "lstm",
                                     output_last_state = T,
                                     config = "one-to-one")

symbol.infer <- rnn.graph.unroll(seq_len = 1,
                                 num_rnn_layer = 2, 
                                 num_hidden = 96,
                                 input_size = vocab,
                                 num_embed = 64, 
                                 num_decode = vocab,
                                 masking = F, 
                                 loss_output = "softmax",
                                 dropout = 0.2, 
                                 ignore_label = -1,
                                 cell_type = "lstm",
                                 output_last_state = T,
                                 config = "one-to-one")
```


### Inference with most likely term

Here the predictions are performed by picking the character whose associated probablility is the highest. 

```{r, echo=TRUE, eval=TRUE}
model <- mx.model.load(prefix = "../models/model_one_to_one_lstm_cpu", iteration = 1)

predict <- numeric()
data = mx.nd.array(matrix(infer_split))

infer <- mxnet:::mx.rnn.infer.unroll(data = data, 
                             symbol = symbol.infer.ini,
                             num_hidden = 96,
                             arg.params = model$arg.params,
                             aux.params = model$aux.params,
                             init_states = NULL,
                             ctx = ctx)

pred_prob <- mx.nd.slice.axis(infer[[1]], axis=0, begin = infer_length-1, end = infer_length)
pred <- mx.nd.argmax(data = pred_prob, axis = 1, keepdims = T)
predict <- c(predict, as.numeric(as.array(pred)))

for (i in 1:100) {
  
  infer <- mxnet:::mx.rnn.infer.unroll(data = pred, 
                               symbol = symbol.infer,
                               num_hidden = 96,
                               arg.params = model$arg.params,
                               aux.params = model$aux.params,
                               init_states = infer[-1],
                               ctx = ctx)
  
  pred <- mx.nd.argmax(data = infer$loss_output, axis = 1, keepdims = T)
  predict <- c(predict, as.numeric(as.array(pred)))
  
}

predict_txt <- paste0(rev_dic[as.character(predict)], collapse = "")
predict_txt_tot <- paste0(infer_raw, predict_txt, collapse = "")
```

Generated sequence: `r predict_txt_tot`

Key ideas appear somewhat overemphasized. 

### Inference from random sample

Noise is now inserted in the predictions by sampling each character based on their modeled probability. 

```{r, echo=TRUE, eval=TRUE}
set.seed(44)

infer_raw <- c("The United States are")
infer_split <- dic[strsplit(infer_raw, '') %>% unlist]
infer_length <- length(infer_split)

predict <- numeric()

infer <- mxnet:::mx.rnn.infer.unroll(data = data, 
                             symbol = symbol.infer.ini,
                             num_hidden = 96,
                             arg.params = model$arg.params,
                             aux.params = model$aux.params,
                             init_states = NULL,
                             ctx = ctx)

pred_prob <- as.numeric(as.array(mx.nd.slice.axis(
  infer[[1]], axis=0, begin = infer_length-1, end = infer_length)))
pred <- sample(length(pred_prob), prob = pred_prob, size = 1) - 1
predict <- c(predict, pred)

for (i in 1:100) {
  
  infer <- mxnet:::mx.rnn.infer.unroll(data = mx.nd.array(array(pred, dim = c(1,1))), 
                               symbol = symbol.infer,
                               num_hidden = 96,
                               arg.params = model$arg.params,
                               aux.params = model$aux.params,
                               init_states = infer[-1],
                               ctx = ctx)
  
  pred_prob <- as.numeric(as.array(infer[[1]]))
  pred <- sample(length(pred_prob), prob = pred_prob, size = 1, replace = T) - 1
  predict <- c(predict, pred)
}

predict_txt <- paste0(rev_dic[as.character(predict)], collapse = "")
predict_txt_tot <- paste0(infer_raw, predict_txt, collapse = "")
```

Generated sequence: `r predict_txt_tot`

Now we get a more alembicated political speech.

